Effect of quality-adjusted production function inputs

### 2015-05-27 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Now putting all CD data on one graph for resids.

<<Fitted_and_Residuals>>=
fitted_and_resid_data <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), f = function(mod, id){
  df <- select(mod$data, Year, iGDP, Country)
  # Trim fitted to be same length as df to eliminate problems with possible NAs at end of fitted
  df$fitted <- yhat(mod)[c(1:length(df$Year))]
  df$Residuals <- resid(mod)[c(1:length(df$Year))]
  # Parse the id to find out whether this is quality-adjusted data
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  return(df)  
}) %>% do.call(rbind, .)

# Get the unadjusted, without-energy residuals and store them in a data frame
r_unadj_noE <- filter(fitted_and_resid_data, flavor=="Unadjusted" & energy=="Without energy") %>% 
               select(Year, Country, Residuals, model)
colnames(r_unadj_noE)[colnames(r_unadj_noE)=="Residuals"] <- "r_unadj_noE"
# Merge to get the unadjusted, without-energy values next to each country, year, and model
fitted_and_resid_data <- merge(fitted_and_resid_data, r_unadj_noE)
# Create a column containing ∆r, the difference between a residual value
# and the residual value in the same year for the unadjusted, no-energy case
fitted_and_resid_data[["Dr"]] <- abs(fitted_and_resid_data[["Residuals"]]) - abs(fitted_and_resid_data[["r_unadj_noE"]])

# Relevel factors before plotting
fitted_and_resid_data$Country <- relevelFactor(as.factor(fitted_and_resid_data$Country), c("PT", "UK"))
fitted_and_resid_data$flavor <- relevelFactor(as.factor(fitted_and_resid_data$flavor), c("Unadjusted", "Quality-adjusted"))
fitted_and_resid_data$energy <- relevelFactor(as.factor(fitted_and_resid_data$energy), c("Without energy", "With energy"))
@

@@

        } else {
          stop(paste("Unknown model", model))
        }
        # Add to the list of data frames
        dframes[[length(dframes)+1]] <- df
      }
    }
  }      
}
solowResiduals <- do.call(plyr::rbind.fill, dframes)

# Grab the unadjusted, no-energy Solow residuals and create a column in solowResiduals with them. 
# Using the hint at 
# http://stackoverflow.com/questions/15059076/
# r-how-to-call-apply-like-function-on-each-row-of-dataframe-with-multiple-argum
sr_unadj_noE_col <- apply(solowResiduals[ , c("Country", "model", "energy", "Year", "flavor")], 1, function(x){
    row <- as.data.frame(filter(solowResiduals, 
                                Country==x["Country"], model==x["model"], energy=="Without energy",
                                Year==x["Year"], flavor=="Unadjusted"))
    return(row[["lambda_fd"]])
  })
solowResiduals[["sr_unadj_noE"]] <- do.call(rbind, as.list(sr_unadj_noE_col))
# Create a column containing ∆lambda_df, the difference between a lambda_df value
# and the lambda_df value in the same year for the unadjusted, no-energy case
solowResiduals[["Dlambda_fd"]] <- abs(solowResiduals[["lambda_fd"]]) - abs(solowResiduals[["sr_unadj_noE"]])

# Adjust levels before creating graphs.
solowResiduals$flavor <- relevelFactor(as.factor(solowResiduals$flavor), c("Unadjusted", "Quality-adjusted"))
solowResiduals$energy <- relevelFactor(as.factor(solowResiduals$energy), c("Without energy", "With energy"))
@

@@

<<SR and R trajectory graph functions>>=
residualsGraph <- function(data){
  plot <- ggplot(data=data) + 
    geom_point(mapping=aes(x=Dr, y=Dlambda_fd)) + 
    xlab(expression(Delta*r)) +
    ylab(expression(Delta*lambda)) +
    xy_theme()
  return(plot)
}
traj.data <- merge(solowResiduals[c("Country", "Year", "model", "flavor", "energy", "nest", "Dlambda_fd")], 
                   fitted_and_resid_data[c("Country", "Year", "model", "flavor", "energy", "nest", "Dr")])
traj.data.means <- traj.data %>% 
  group_by(model, Country, flavor, energy, nest) %>%
  summarise(Dlambda_fd.mean = mean(Dlambda_fd, na.rm=TRUE), Dr.mean = mean(Dr))
@

@@

<<Solow_residual_trajectories, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual trajectories for the Cobb-Douglas model.">>=
residualsGraph(subset(traj.data, model=="CD" & Country=="UK" & energy=="With energy" & flavor=="Unadjusted"))
@

>>>>

<<Fitted_and_Residuals>>=
temp_data <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), f = function(mod, id){
  df <- select(mod$data, Year, iGDP, Country)
  # Trim fitted to be same length as df to eliminate problems with possible NAs at end of fitted
  df$fitted <- yhat(mod)[c(1:length(df$Year))]
  df$Residuals <- resid(mod)[c(1:length(df$Year))]
  # Parse the id to find out whether this is quality-adjusted data
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  return(df)  
}) %>% do.call(rbind, .)

# Get the unadjusted, without-energy residuals and store them in a data frame
r_unadj_noE <- filter(temp_data, flavor=="Unadjusted" & energy=="Without energy") %>% 
               select(Year, Country, Residuals, model)
colnames(r_unadj_noE)[colnames(r_unadj_noE)=="Residuals"] <- "r_unadj_noE"
# Merge to get the unadjusted, without-energy values next to each country, year, and model
fitted_and_resid_data <- merge(temp_data, r_unadj_noE)
# Create a column containing ∆r, the difference between a residual value
# and the residual value in the same year for the unadjusted, no-energy case
fitted_and_resid_data[["Dr"]] <- abs(fitted_and_resid_data[["Residuals"]]) - abs(fitted_and_resid_data[["r_unadj_noE"]])

# Relevel factors before plotting
fitted_and_resid_data$Country <- relevelFactor(as.factor(fitted_and_resid_data$Country), c("PT", "UK"))
fitted_and_resid_data$flavor <- relevelFactor(as.factor(fitted_and_resid_data$flavor), c("Unadjusted", "Quality-adjusted"))
fitted_and_resid_data$energy <- relevelFactor(as.factor(fitted_and_resid_data$energy), c("Without energy", "With energy"))
fitted_and_resid_data$nest <- relevelFactor(as.factor(fitted_and_resid_data$nest), c("kl", "kle", "lek", "ekl"))

# Make data viewing easier
fitted_and_resid_data <- arrange(fitted_and_resid_data, Country, model, flavor, energy, nest)
@

@@

        } else {
          stop(paste("Unknown model", model))
        }
        # Add to the list of data frames
        dframes[[length(dframes)+1]] <- df
      }
    }
  }      
}
temp_data <- do.call(plyr::rbind.fill, dframes)

# Get the unadjusted, without-energy Solow residuals and store them in a data frame
sr_unadj_noE <- filter(temp_data, flavor=="Unadjusted" & energy=="Without energy") %>% 
                select(Year, Country, lambda_fd, model)
colnames(sr_unadj_noE)[colnames(sr_unadj_noE)=="lambda_fd"] <- "sr_unadj_noE"
# Merge to get the unadjusted, without-energy values next to each country, year, and model
solowResiduals <- merge(temp_data, sr_unadj_noE)

# Create a column containing ∆lambda_df, the difference between a lambda_df value
# and the lambda_df value in the same year for the unadjusted, no-energy case
solowResiduals[["Dlambda_fd"]] <- abs(solowResiduals[["lambda_fd"]]) - abs(solowResiduals[["sr_unadj_noE"]])

# Relevel factors before plotting
solowResiduals$Country <- relevelFactor(as.factor(solowResiduals$Country), c("PT", "UK"))
solowResiduals$flavor <- relevelFactor(as.factor(solowResiduals$flavor), c("Unadjusted", "Quality-adjusted"))
solowResiduals$energy <- relevelFactor(as.factor(solowResiduals$energy), c("Without energy", "With energy"))
solowResiduals$nest <- relevelFactor(as.factor(solowResiduals$nest), c("kl", "kle", "lek", "ekl"))

# Make data viewing easier
solowResiduals <- arrange(solowResiduals, Country, model, flavor, energy, nest)
@

@@

<<SR and R trajectory data>>=
traj.data <- merge(solowResiduals[c("Country", "Year", "model", "flavor", "energy", "nest", "Dlambda_fd")], 
                   fitted_and_resid_data[c("Country", "Year", "model", "flavor", "energy", "nest", "Dr")]) %>%
                   arrange(Country, model, flavor, energy, nest)
traj.data.means <- traj.data %>% 
  group_by(model, Country, flavor, energy, nest) %>%
  summarise(Dlambda_fd.mean = mean(Dlambda_fd, na.rm=TRUE), Dr.mean = mean(Dr))
@

<<SR and R trajectory plot function>>=
residualsTrajGraph <- function(data){
  plot <- ggplot(data=data) + 
    geom_point(mapping=aes(x=Dr, y=Dlambda_fd, color=energy)) + 
    facet_grid(Country ~ flavor) +
    xlab(expression(Delta*r)) +
    ylab(expression(Delta*lambda)) +
    xy_theme()
  return(plot)
}
@

@@

<<Solow_residual_trajectories, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual trajectories for the Cobb-Douglas model.">>=
residualsTrajGraph(filter(traj.data, model=="CD"))
@
